A hierarchical model to evaluate pest treatments from prevalence and intensity data

### 2023-09-08 Review/response_strategy.Rmd

---
title: "R Notebook"
output: 
  html_notebook:
    toc: yes
    toc_float: yes
    code-fold: yes
---

```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 300px;
}

pre, code {white-space:pre !important; overflow-x:scroll !important}

.question {
  text-indent: 5em;
  text-align: justify;
}

.indented_response {
  text-indent: 5em;
  color: #000080;
  text-align: justify;
  margin-bottom: 30px;
}

.response {
  color: #000080;
  text-align: justify;
  margin-bottom: 30px;
  margin-left: 40px
}

.rouge {
  color: red;
  text-align: justify;
  margin-bottom: 30px;
  margin-left: 40px
} 
```

```{r, results = FALSE, cache = TRUE}
library(ggplot2);   library(dplyr);   library(rjags);   library(kableExtra)
```

# Reviewer 1

::: question
In the manuscript 'A hierarchical model to evaluate pest treatments from prevalence and intensity data', Favrot & Makovski present a new and welcome (Bayesian) method for the (meta-)analysis of countable 'infection' units on individual hosts, where the recorded variable may be either the proportion of 'infected' hosts (prevalence) or the mean number of infection units per individual host (intensity), or both quantities. Then, a simulation study is used to demonstrate that intensity data provide more accurate estimates than prevalence data (especially for abundant 'infections'), and to assess the accuracy of the estimates depending on the number of trials. The provided code and convergence seems good. 
:::

::: question
Before starting my comments, I wish to apologize deeply for the inconvenience caused to the authors and editor by the time taken to review the manuscript. Besides, I did not find any line numbering; thus, my comment below only vaguely refer to paragraph or page numbers.
:::

::: indented_response
Ajouter des numéros de lignes dans mes fichiers.
:::

::: question
A general comment : the article focusses on aphid abundance on plants sampled during field trials designed to compare the efficacy of various pest control options, but the framework is more generic (and might be used e.g. in animal/human epidemiology) and the authors may highlight this point in the abstract, introduction and/or discussion.
:::

::: indented_response         
Ajouter une phrase dans la conclusion pour souligner ce point.
:::

## Major comments

::: question
My only methodological comment relates to the choice of a Poisson distribution for W, which is the number of aphids on each plant. Because aphids land and initiate exponentially growing clonal populations at different times on different plants of the same trial x treatment x block x time combination, W should have an overdispersed Poisson distribution (e.g., negative binomial), instead of a Poisson distribution (eq. 1). Thus, Y (the sum of N independent W variables) should not have a Poisson distribution (eq. 3). It would be interesting to test whether an overdispersed Poisson distribution is better than a Poisson distribution for fitting the unaggregated data on the number of aphids on each of the N_i (i.e. 10 here) sampled plants (gaining statistical power from the multiple trials). Although the authors state (in part 2.1) that such data are not available to the authors, these raw data are classically recorded and might be provided by the unspecified colleagues who conducted the field trials (by the way, the institutions that performed the field trials should be mentioned more explicitly).
:::

::: indented_response         
Ce commentaire remet en question le modèle que l'on utilise et propose d'en utiliser un autre : la négative binomiale. La négative binomiale permet de gérer la surdispersion. Ce commentaire nous a fait remarquer que nous n'avions pas vérifié l'ajustement aux données de notre modèle. La première étape sera donc de faire cette vérification. On la fera en appliquant la méthode du "posterior predictive checking", qui consiste, pour chaque observation, à calculer la probabilité que le modèle propose une valeur plus extrême que celle de l'observation considérée. On fera cette vérification uniquement sur les données réelles. La seconde étape sera de coder le modèle reposant sur la loi négative binomiale et de le comparer au modèle de Poisson en utilisant le posterior predictive checking : on fera un scatter plot des proba à posteriori de dépasser les valeurs extrêmes selon les deux modèles (probas d'un modèle en abscisse et probas de l'autre en ordonné).

Notre stratégie de réponse dépendra des résultats de ces deux premières étapes.
:::

[paragraphe biblio pour justifier notre choix]

Nous avons implémenté le modèle négatif binomial sur les données réelles et nous avons comparé son ajustement aux données par rapport au modèle de Poisson, en utilisant la méthode probability posterior checking **[ref]**.

Le modèle est donné par

$$
    W_{ijkts}\sim\mathcal{NB}\big(r,\ \lambda_{ijkt}\big)
$$

Where $\lambda_{ijkt}$ is the mean of $W_{ijkts}$ and $r$ is the overdispersion parameter. The model on $\lambda_{ijkt}$ is the same as the one of the Poisson model, but without the term $\epsilon_{ijkt}$ :

$\begin{align}&\log\ \lambda_{ijkt} = \alpha_0 + \beta_{0i} + \gamma_{0j}  + (\alpha_1 + \gamma_{1j})\ X_t + u_{ij}\label{eq:test}\tag{1}\end{align}$

This settings leads to the following distribution for $Y$ and $Z$ :

$$Y_{ijkt} \sim\mathcal{NB}\big(N_ir,\ N_i\lambda_{ijkt}\big)\ \hspace{1cm} \ Z_{ijkt} \sim \mathcal{B}\big(N_{i},\ \pi_{ijkt}\big)$$

where $\pi_{ijkt} = 1-\big(r\ /\ (r + \lambda_{ijkt})\big) ^ r$ is the probability of $W_{ijkts}$ beeing postitive.

La figure  ci-dessous montre les résultats obtenus pour les deux modèles. On voit que l'ajustement au donné est meilleur avec le modèle de poisson.

![Posterior predictive check](Posterior_predictive_check.png)

![Observed vs predicted values](Observed_vs_predicted_values.png){#fig-Observed_vs_predicted_values_real_data}

<br>
Nous avons aussi vérifié par simulation que notre modèle était pertinent. Nous avons simulé selon le modèle de la négative binomiale puis nous avons inférer avec 2 modèles : 
  - le modèle négative binomiale
  - notre modèle de poisson surdispersé

Le code est donné ci-dessous. Les résultats montre qu'encore une fois le modèle de poisson s'ajuste mieux aux données même si celles si ont été simulé avec le modèle de la négative binomiale.

**Simulation des données**

```{r, cache=TRUE}

source(file = "simu_data_NegB.R")

I = 30
K = 4   # number of blocks
S = 10   # number of beets in the sample

r = 15   # overdispersion parameter

alpha_0 = 0.5
alpha_1 = 0.16

gamma_0 = c(0, - 0.13, - 1.13, - 1.24);  
gamma_1 = c(0, 0.24, - 0.14, - 0.15);

X_t = c(6, 12)

sigma2 = 1.87;  
chi2 = 0.27

set.seed(seed = 1)

data_simu = simu_data_NegB(I = I, K = K, S = S, r = r, alpha_0 = alpha_0, alpha_1 = alpha_1, 
                      gamma_0 = gamma_0, gamma_1 = gamma_1, X_t = X_t, sigma2 = sigma2, chi2 = chi2)

ggplot(data_simu) +
    geom_histogram(aes(x = Y), binwidth = 5)
```

```{r, cache = TRUE}
# Making the two scenarios [100% Z - 0% Y] and [0% Z - 100% Y]
dataZ = data_simu %>% mutate(Y = NA)
dataY = data_simu %>% mutate(Z = NA)
```

**Ecriture des modèles**

```{r, cache = TRUE}
# Negative binomial model

negative_binomial_model = "
model {

    ###################### Likelyhood ######################
    for (i in 1:Q){
        Y[i] ~ dnegbin(p[i], N[i] * r)
        p[i] <- r / (r + lambda[i])

        Z[i] ~ dbinom(pi[i], N[i])
        pi[i] = 1 - p[i] ^ r

        log(lambda[i]) = beta0[TRIAL[i]] + gamma0[TREATMENT[i]] + (alpha1 + gamma1[TREATMENT[i]]) * TIME[i] + u[TRIAL_TREATMENT[i]]

        pred[i] ~ dnegbin(p[i], N[i] * r)            
    }

    for (j in 1:K){
        beta0[j] ~ dnorm(alpha0, invsigma)
    }

    for (c in 1:M){
        u[c] ~ dnorm(0, invchi)
    }

    gamma0[1] = 0
    gamma1[1] = 0

    ######################## Priors ########################
    for (s in 2:L){
        gamma0[s] ~ dnorm(0, 0.001)
        gamma1[s] ~ dnorm(0, 0.001)
    }

    alpha0 ~ dnorm(0, 0.001)
    alpha1 ~ dnorm(0, 0.001)
    sigma ~ dunif(0, 10)
    chi ~ dunif(0, 10)
    r ~ dunif(0, 50)

    ################### Derived Quantities #################
    invsigma = pow(sigma, -2)
    invchi = pow(chi, -2)
}
"

writeLines(negative_binomial_model, con  =  "negative_binomial_model.txt")

# Poisson model

poisson_model = "
model {

    ###################### Likelyhood ######################
    for (i in 1:Q){
        Y[i] ~ dpois(N[i] * lambda[i])
        Z[i] ~ dbinom(pi[i], N[i])
        log(lambda[i]) = beta0[TRIAL[i]] + gamma0[TREATMENT[i]] + (alpha1 + gamma1[TREATMENT[i]]) * TIME[i] + u[TRIAL_TREATMENT[i]] + epsi[i]
        pi[i] = 1 - exp(- lambda[i])
        epsi[i] ~ dnorm(0, inveta)
        pred[i] ~ dpois(N[i] * lambda[i])
    }

    for (j in 1:K){
        beta0[j] ~ dnorm(alpha0, invsigma)
    }

    for (c in 1:M){
        u[c] ~ dnorm(0, invchi)
    }

    gamma0[1] = 0
    gamma1[1] = 0

    ######################## Priors ########################
    for (s in 2:L){
        gamma0[s] ~ dnorm(0, 0.001)
        gamma1[s] ~ dnorm(0, 0.001)
    }

    alpha0 ~ dnorm(0, 0.001)
    alpha1 ~ dnorm(0, 0.001)
    sigma ~ dunif(0, 10)
    chi ~ dunif(0, 10)
    eta ~ dunif(0, 10)

    ################### Derived Quantities #################
    invsigma = pow(sigma, -2)
    invchi = pow(chi, -2)
    inveta = pow(eta, -2)
}
"

writeLines(poisson_model, con  =  "poisson_model.txt")
```

**Inference**

```{r, results = FALSE, cache = TRUE}
# Function to make inference from the data ##########################################################################################
finf <- function(model, df, nadapt, niter){

    Y = df$Y;   Q = length(Y);   N = df$N;   Z = df$Z

    TRIAL = as.numeric(as.factor(as.character(df$trial)));   TREATMENT = as.numeric(as.factor(as.character(df$treatment)));

    TIME = as.numeric(df$X_t);   TRIAL_TREATMENT = as.numeric(as.factor(as.character(df$trial_treatment)));

    K = length(unique(TRIAL));   L = length(unique(TREATMENT));   M = length(unique(TRIAL_TREATMENT));

    # Inférence
    model <- jags.model(file = model, 
                        data = list("Y" = Y, "Z" = Z, "Q" =  Q, "TRIAL" = TRIAL, "TREATMENT" = TREATMENT, "TIME" = TIME, 
                                    "TRIAL_TREATMENT" = TRIAL_TREATMENT, "K" = K, "L" = L, "M" = M, "N" = N), 
                        n.chains = 2, 
                        n.adapt = nadapt)

    samples <- coda.samples(model, variable.names = c("pred"), n.iter = niter, thin = 10)

    return(samples)
}

nadapt = 2000;   niter = 4000
```

```{r, results = FALSE, cache = TRUE}
# Inference with the negative binomiale model, using Intensity
resY_posterior_checking_negative_binomiale = finf(model = "negative_binomial_model.txt", dataY, nadapt = nadapt, niter = niter)

# Inference with the Poisson model, using intensity
resY_posterior_checking_poisson = finf(model = "poisson_model.txt", dataY, nadapt = nadapt, niter = niter)
```

**Comparaison des résultats**

```{r, cache = TRUE}
df_prediction = rbind(

        # negative binomial samples
        cbind(t(resY_posterior_checking_negative_binomiale[[1]]), t(resY_posterior_checking_negative_binomiale[[2]])) %>% 
            as.data.frame %>% mutate(model = "Negative binomial model"),

        # poisson samples
        cbind(t(resY_posterior_checking_poisson[[1]]), t(resY_posterior_checking_poisson[[2]])) %>% 
            as.data.frame %>% mutate(model = "Poisson model")
    ) %>%

    mutate(observation = rep(data_simu$Y, 2)) %>% # 2 for the two models
    mutate(prediction = rowMeans(across(starts_with("V")))) %>%
    mutate(error = (prediction - observation) ^ 2)

ggplot(df_prediction) +
    geom_point(aes(x = observation, y = prediction)) +
    facet_wrap(~ model)
```

```{r, cache = TRUE}
df_prediction %>%
    summarise(MSE = sum(error), .by = model) %>% kbl() %>% kable_styling()
```

```{r, cache = TRUE}
df_prediction %>%
    summarise(MSE = sum(error), .by = model)
```

**MSE sur l'estimation des paramètres**

*En fréquentiste pour réduire le temps de calcul*

```{r mse_simulations, cache = TRUE}
library(lme4)

I_vec = c(30)
seeds = c(1 : 10)

res = NULL

for(I in I_vec){
    for(seed in seeds){

        set.seed(seed)

        data_simu = simu_data_NegB(I = I, K = K, S = S, r = r, alpha_0 = alpha_0, alpha_1 = alpha_1, 
                              gamma_0 = gamma_0, gamma_1 = gamma_1, X_t = X_t, sigma2 = sigma2, chi2 = chi2)

        res_negB = glmer.nb(Y ~ (1|trial) + treatment * X_t + (1|trial_treatment), 
                            data = data_simu)

        res_Poisson = glmer(Y ~ (1|trial) + treatment * X_t + (1|trial_treatment) + (1|ligne), 
                            data = data_simu,
                            family = "poisson")

        MSE_poisson = sum((c(alpha_0 + log(S), gamma_0[- 1], alpha_1, gamma_1[- 1]) - summary(res_Poisson)$coefficients[, 1]) ^ 2)

        MSE_NegB = sum((c(alpha_0 + log(S), gamma_0[- 1], alpha_1, gamma_1[- 1]) - summary(res_negB)$coefficients[, 1]) ^ 2)

        res = rbind(res, data.frame(I = I,
                                    seed = seed,
                                    Model = c("Poisson with OLRE", "Negative binomiale"),
                                    MSE = c(MSE_poisson, MSE_NegB),
                                    Message = c(summary(res_Poisson)$optinfo$message,
                                                summary(res_negB)$optinfo$message),
                                    Warning = c(summary(res_Poisson)$optinfo$warnings %>% length,
                                                summary(res_negB)$optinfo$warnings %>% length)))
    }
}
```

```{r}
ggplot(res) +
    geom_boxplot(aes(y = MSE, x = factor(I), fill = Model))
```

## Minor comments

My other comments are minor and mainly concern typos:

- &#10060; p.3 l.2: 'indeed more time-dependent'

::: response
Ce commentaire fait référence à cette phrase (dans l'introduction) : "measuring intensity is indeed often more time demanding than measuring prevalence, in particular when the pest individuals are small, numerous, and difficult to detect." Je ne comprends pas ce commentaire.
:::

- &#9989; p.3: 'negative impact on a wide range', 'this family of pesticides has been banned', 'some trials may report'

  - &#9989; 'negative impact on a wide range'

  ::: response
  fait référence à la phrase : "Recently, neonicotinoids were recognized as presenting high risks for the environment with a negative impact a wide range of non-target organisms"

  =\> on a oublié le "on", corriger
  :::

  - &#9989; 'this family of pesticides has been banned'

  ::: response
  fait référence à la phrase : "and this familly of pesticides had been banned in several European countries"

  =\> erreur de temps, changer had par has
  :::

  - &#9989; 'some trials may report'

  ::: response
  fait référence à la phrase : "Consequently, for a given pest treatment, some of trials may report prevalence data while others report intensity data"

  =\> le "of" est en trop, corriger
  :::

- &#9989; p.3: 'both types of data', replace 'will' with 'may' (twice)

  - &#9989; ‘both types of data’

  ::: response
  fait référence à la phrase : "while others report intensity data or both type of data"

  => il manque un “s” à “type”, corriger
  :::

  - &#9989; replace ‘will’ with ‘may’ (twice)

  ::: response
  les deux will en questions sont dans la phrase suivante : "It is unclear whether the simultaneous analysis of prevalence and intensity data will increase the accuracy of the estimated treatment efficacy compared to the use of a single type of data, and whether this will increase the probability of identifying the most efficient treatments" =\> faire le remplacement
  :::

- p.3-4: please rephrase this sentence for clarity

::: response
Le reviewer fait référence à cette phrase : "Based on simulations, we then quantify the reduction of mean absolute errors in the estimated treatment efficacies induced by the combination of prevalence and incidence data compared to the use of a single type of data."

Tout d'abord on avait une erreur dans cette phrase : ce n'est pas incidence mais intensity (incidence est synonyme de prévalence).

**Proposition de reformulation** : "Based on simulations, we then quantify the reduction of mean absolute errors in the estimated treatment efficacies induced by using both prevalence and intensity data, compared to the use of exclusively either prevalence or intensity data."
:::

- &#9989; 2.1: 'the Netherlands', 'Intensity and prevalence were monitored at different times after treatments'

  - &#9989; 'the Netherlands'

  ::: response
  fait référence à la phrase : "Data were collected in 32 field trials conducted in France, Belgium and Netherlands to compare several treatments"

  =\> oubli du "the", corriger
  :::

  - &#9989; 'Intensity and prevalence were monitored at different times after treatments'

  ::: response
  fait référence à la phrase : "Intensity and prevalence were monitored at different dates during the growing season."

  =\> remplacer "dates during the growing season" par "times after treatments"
  :::

- &#9989; Fig.1 legend: last 2 sentences are not necessary (obvious or redundant with the figure)

::: response
=\> supprimer ces deux phrases
:::

- &#9989; p.6: 'assuming the Ws are independent', 'quantities used in the model'

  - &#9989; ‘assuming the Ws are independent’

  ::: response
  fait référence à la phrase : "Then, assuming the Ws independent, we have:"

  =\> rajouter "are"
  :::

  - &#9989; ‘quantities used in the model’

  ::: response
  fait référence à la phrase : "The different quantities used by the model are defined in Table 1."

  =\> changer "by" par "in"
  :::

### 2023-09-08 Article/Article.qmd
change in notation for scenarios

::: {.callout-note}
The following code presents the inference on an extract of the real 
dataset, which includes both trials of figure 1B (2020 - B1A97 ; 2020 - 
u1CwE). It is  a demo for the "50% Y / 50% Z" scenario and we set here the 
number of adaptation and iteration to 2000 in order to reduce computation 
time.
:::

@@

\
In practice, it is common that only $Y$ or $Z$ data are available in some of the trials. In this case, the resulting dataset includes observations of $Y$ in some trials and observations of $Z$ in others. The dataset may even include one type of observations only, either $Y$ or $Z$, in all trials. Here, we define four scenarios with contrasted levels of $Y$ and $Z$ availability in order to evaluate the consequences of using different types of datasets. We consider four datasets including trials with observations of $Y$, with observations of $Z$, or with both types of observation in different proportions (Table 2). The dataset "100% Y" includes Y data collected in the 32 trials. The dataset "50%Y" includes Y data collected in the 17 trials for which no Z observation is available. The dataset "50% $Z$" includes the Z data collected in the 15 trials for which Z observations are available. The dataset "50% $Y$ - 50% $Z$" includes $Y$ data collected in 17 trials and $Z$ data collected in the other 15 trials. The later dataset does include any trial reporting both $Y$ and $Z$ data.

@@

| Type of dataset | Description |
|---------|:-----|
| 100% Y      | Y observations available in the 32 trials   |
| 50% Y     | Y observations available in 17 trials and no Z  | 
| 50% Z       | Z observations available in 15 trials and no Y    | 
| 50% Y / 50% Z       | Y observations available in 17 trials and Z observations available in the other 15 trials   |

@@

We define three numbers of trials, equal to 20, 40 and 80, successively. For data simulations, the model parameters are set equal to those estimated from the real dataset "100% Y" defined in Table 2 (posterior means). For each number of trials, we generate virtual data from the model (@eq-model_W - @eq-model_Z) and estimate the model parameters, according to the following procedure:

@@

In order to determine whether the difference of performance obtained with the types of data $Y$ and $Z$ depends on the pest abundance, we perform an additional series of simulations with three values for the model parameter $\alpha_0$, equal to $-1,\ 1$ and $2$, successively. These three values of $\alpha_0$ define three contrasted levels of pest abundance (the higher $\alpha_0$, the higher the abundance). We used the procedure outlined above considering only two scenarios of data availability, namely "100% Y" and "100% Z". This procedure is implemented with each value of $\alpha_0$ in turn. The results are used to compare the model performances using either $Y$ or $Z$ for parameter estimation, depending of the pest abundance specified by $\alpha_0$.

|Type of dataset| Description|
|---|---|
|100% W | W observations available in all the trials| 
|100% Y | Y observations available in all the trials|
|100% Z | Z observations available in all the trials|
|100% Y - 100 Z | Y and Z observations available in all the trials|
|50% Y | Y observations available in half of the trials|
|50% Z | Z observations available in half of the trials|
|50% Y / 50% Z | Y observations available in half of the trials and Z observations available in the other half of the trials|
|100% Z / 50% Y | Y observations available in half of the trials and Z observations available in all the trials|

@@

Table 4 gives a summary of the posterior distributions of the model parameters and treatment efficacies obtained for the "100% Y" scenario. The significantly positive value of $\alpha_1$ indicates that the aphid numbers tend to increase with time in untreated plots. The relatively high value of $\sigma_0$ (posterior mean equal to 1.87) reveals a strong variability in aphid numbers between trials. The posterior mean value of $\chi$ (0.27) suggests that the treatment efficacy varies across trials. The $\gamma_0$ parameter is negative for all three treatments, indicating a negative effect of the treatments on the aphid numbers at the time of pesticide spray. The Movento and Teppeki treatments have a similar effect with a posterior mean for $\gamma_0$ equal to -1.13 and -1.24, and a standard deviation equal to 0.12 and 0.11, respectively. The effect of treatment Mavrik Jet is weaker as its posterior mean for $\gamma_0$ is equal to -0.13 and its standard deviation is equal to 0.16. The $\gamma_1$ posterior means are negative for Movento and Teppeki (-0.14 and -0.15), suggesting that the effect of these treatments tend to increase with time, but the posterior mean value is positive for Mavrik Jet (= 0.24), suggesting that the effect of this treatment may decrease with time. However, the 95% credibility intervals of $\gamma_1$ include zero and these parameters are not very accurately estimated.

```{r figure2, cache = TRUE, out.width="100%", fig.height = 7, echo = FALSE}
#| label: fig-two
#| fig-cap: "Model convergence - Markov chain for the model parameters and treatment efficacies, in the scenarios '50% Y' (**A**),  '50% Y / 50% Z' (**B**), '100% Y' (**C**) and '50% Z' (**D**). The x-axis presents the iteration number and the y-axis presents the sampled value."

@@

```{r figure3, cache = TRUE, out.width="100%", fig.height = 7, echo = FALSE}
#| label: fig-three
#| fig-cap: "Model convergence - Gelman-Rubin statistics for the parameters of the model (@eq-model_W - @eq-model_Z) and for treatment efficacies (@eq-Efficacy), according to the scenarios 50% Y (**A**),  50% Y / 50% Z (**B**), 100% Y (**C**) and 50% Z (**D**). The x-axis presents the iteration number and the y-axis presents the Gelman-Rubin statistic."

@@

@fig-four presents the posterior means and the 95% credibility intervals of treatment efficacies at 6 days (A) and 12 days (B) after pesticide spray, for the "100% Y", "50% Y / 50% Z", "50% Y" and "50% Z" scenarios. Different scenarios are indicated by different colors. The x-axis presents the efficacy and the y-axis presents the treatments. Overall, the results obtained are consistent across scenarios; Teppeki and Movento show higher mean efficacies than Mavrik Jet, and the credibility intervals are narrower for Teppeki and Movento than for Mavrik Jet in all scenarios. The credibility interval of the "100% Y" scenario is narrower than that of the "50% Y / 50% Z" scenario, which is itself narrower than that of the "50% Y" and "50% Z" scenarios. Table 5 quantifies the differences in the sizes of the credibility intervals and present the reduction percentage in the credibility interval sizes of the "100% Y", "50% Y / 50% Z" and "50% Z" scenarios, compared to the "50% Y" scenario, for all the treatment efficacies at 6 days (third column) and at 12 days (fourth column) after pesticide spray. A positive (negative) value corresponds to an increase (decrease) of the credibility interval size compared to "50% Y". For example, the credibility interval size obtained for the efficacy of Movento at 6 days is 40.5% smaller with the "100% Y" scenario than with the "50% Y" scenario. Overall, the credibility interval sizes obtained with the "100% Y" scenario are 25 to 45% smaller than those obtained with the "50% Y" scenario. This indicates that the accuracy of the estimated treatment efficacies is improved by doubling the number of Y observations. Results show that the credibility intervals are often larger with "50% Z" than with "50% Y", suggesting that better esimtations are obtained using Y than Z when only a limited number of data is available, at least in this case study. Interestingly, the sizes of the credibility intervals are smaller by about 25% with "50% Y / 50% Z" compared to "50% Y", showing that the combination of Y and Z observations collected in different trials in beneficial and lead to a reduction of the uncertainty in the estimated treatment efficacy. The latter result illustrates how the estimation of treatment efficacy could be improved by combining trials with prevalence data and trials with intensity data.

@@

df_IC = rbind(df_resYhalf %>% mutate(Data = "50% Y"), 
              df_resYZ %>% mutate(Data = "50% Y / 50% Z"),
              df_resY %>% mutate(Data = "100% Y"),
              df_resZhalf %>% mutate(Data = "50% Z"))

@@

df_IC_6 = df_IC %>% mutate(Ordonnee = recode(Insecticide, !!!x)) %>%
                      mutate(Ordonnee = ifelse(Data == "100% Y", Ordonnee + y2, 
                                               ifelse(Data == "50% Y / 50% Z", Ordonnee + y1,
                                                      ifelse(Data == "50% Y", Ordonnee - y1, Ordonnee - y2))),
                             jour = "6")

# calcul du pourcentage de réduction de l'IC par rapport au scenario 50% Y
df_len = df_IC %>% select(Insecticide, name, value, Data) %>% pivot_wider(names_from =  c(name)) %>% as.data.frame %>%
    mutate(IC_length = b_sup - b_inf) %>% arrange(Insecticide)

df_remp = df_len %>% filter(Data == "50% Y") %>% select(Insecticide, IC_length)
x  = df_remp$IC_length;   names(x) = df_remp$Insecticide

@@

df_len6 = df_len %>% rename(Insecticide = Insecticide) %>% select(Insecticide, Data, `For efficacy at 6 days`) %>% 
    filter(Data != "50% Y") %>% arrange(Data) %>% mutate_if(is.numeric, round, digits = 1)

>>>>

::: {.callout-note}
The following code presents the inference on an extract of the real 
dataset, which includes both trials of figure 1B (2020 - B1A97 ; 2020 - 
u1CwE). It is  a demo for the "50% Y - 50% Z" scenario and we set here the 
number of adaptation and iteration to 2000 in order to reduce computation 
time.
:::

@@

\
In practice, it is common that only $Y$ or $Z$ data are available in some of the trials. In this case, the resulting dataset includes observations of $Y$ in some trials and observations of $Z$ in others. The dataset may even include one type of observations only, either $Y$ or $Z$, in all trials. Here, we define four scenarios with contrasted levels of $Y$ and $Z$ availability in order to evaluate the consequences of using different types of datasets. We consider four datasets including trials with observations of $Y$, with observations of $Z$, or with both types of observation in different proportions (Table 2). The dataset "100% Y - 0% Z" includes Y data collected in the 32 trials. The dataset "50%Y" includes Y data collected in the 17 trials for which no Z observation is available. The dataset "50% $Z$" includes the Z data collected in the 15 trials for which Z observations are available. The dataset "50% $Y$ - 50% $Z$" includes $Y$ data collected in 17 trials and $Z$ data collected in the other 15 trials. The later dataset does include any trial reporting both $Y$ and $Z$ data.

@@

| Type of dataset | Description |
|---------|:-----|
| 100% Y - 0% Z      | Y observations available in the 32 trials and no Z   |
| 50% Y - 0% Z     | Y observations available in 17 trials and no Z  | 
| 50% Z - 0% Y       | Z observations available in 15 trials and no Y    | 
| 50% Y - 50% Z       | Y observations available in 17 trials and Z observations available in the other 15 trials   |

@@

We define three numbers of trials, equal to 20, 40 and 80, successively. For data simulations, the model parameters are set equal to those estimated from the real dataset "100% Y - 0% Z" defined in Table 2 (posterior means). For each number of trials, we generate virtual data from the model (@eq-model_W - @eq-model_Z) and estimate the model parameters, according to the following procedure:

@@

In order to determine whether the difference of performance obtained with the types of data $Y$ and $Z$ depends on the pest abundance, we perform an additional series of simulations with three values for the model parameter $\alpha_0$, equal to $-1,\ 1$ and $2$, successively. These three values of $\alpha_0$ define three contrasted levels of pest abundance (the higher $\alpha_0$, the higher the abundance). We used the procedure outlined above considering only two scenarios of data availability, namely "100% Y - 0% Z" and "100% Z - 0% Y". This procedure is implemented with each value of $\alpha_0$ in turn. The results are used to compare the model performances using either $Y$ or $Z$ for parameter estimation, depending of the pest abundance specified by $\alpha_0$.

|Type of dataset| Description|
|---|---|
|100% W | W observations available in all the trials| 
|100% Y - 0% Z | Y observations available in all the trials|
|100% Z - 0% Y | Z observations available in all the trials|
|100% Y - 100 Z | Y and Z observations available in all the trials|
|50% Y - 0% Z | Y observations available in half of the trials|
|50% Z - 0% Y | Z observations available in half of the trials|
|50% Y - 50% Z | Y observations available in half of the trials and Z observations available in the other half of the trials|
|100% Z - 50% Y | Y observations available in half of the trials and Z observations available in all the trials|

@@

Table 4 gives a summary of the posterior distributions of the model parameters and treatment efficacies obtained for the "100% Y - 0% Z" scenario. The significantly positive value of $\alpha_1$ indicates that the aphid numbers tend to increase with time in untreated plots. The relatively high value of $\sigma_0$ (posterior mean equal to 1.87) reveals a strong variability in aphid numbers between trials. The posterior mean value of $\chi$ (0.27) suggests that the treatment efficacy varies across trials. The $\gamma_0$ parameter is negative for all three treatments, indicating a negative effect of the treatments on the aphid numbers at the time of pesticide spray. The Movento and Teppeki treatments have a similar effect with a posterior mean for $\gamma_0$ equal to -1.13 and -1.24, and a standard deviation equal to 0.12 and 0.11, respectively. The effect of treatment Mavrik Jet is weaker as its posterior mean for $\gamma_0$ is equal to -0.13 and its standard deviation is equal to 0.16. The $\gamma_1$ posterior means are negative for Movento and Teppeki (-0.14 and -0.15), suggesting that the effect of these treatments tend to increase with time, but the posterior mean value is positive for Mavrik Jet (= 0.24), suggesting that the effect of this treatment may decrease with time. However, the 95% credibility intervals of $\gamma_1$ include zero and these parameters are not very accurately estimated.

```{r figure2, cache = TRUE, out.width="100%", fig.height = 7, echo = FALSE}
#| label: fig-two
#| fig-cap: "Model convergence - Markov chain for the model parameters and treatment efficacies, in the scenarios '50% Y - 0% Z' (**A**),  '50% Y - 50% Z' (**B**), '100% Y - 0% Z' (**C**) and '50% Z - 0% Y' (**D**). The x-axis presents the iteration number and the y-axis presents the sampled value."

@@

```{r figure3, cache = TRUE, out.width="100%", fig.height = 7, echo = FALSE}
#| label: fig-three
#| fig-cap: "Model convergence - Gelman-Rubin statistics for the parameters of the model (@eq-model_W - @eq-model_Z) and for treatment efficacies (@eq-Efficacy), according to the scenarios '50% Y - 0% Z' (**A**),  '50% Y - 50% Z' (**B**), '100% Y - 0% Z' (**C**) and '50% Z - 0% Y'(**D**). The x-axis presents the iteration number and the y-axis presents the Gelman-Rubin statistic."

@@

@fig-four presents the posterior means and the 95% credibility intervals of treatment efficacies at 6 days (A) and 12 days (B) after pesticide spray, for the "100% Y - 0% Z", "50% Y - 50% Z", "50% Y - 0% Z" and "50% Z - 0% Y" scenarios. Different scenarios are indicated by different colors. The x-axis presents the efficacy and the y-axis presents the treatments. Overall, the results obtained are consistent across scenarios; Teppeki and Movento show higher mean efficacies than Mavrik Jet, and the credibility intervals are narrower for Teppeki and Movento than for Mavrik Jet in all scenarios. The credibility interval of the "100% Y - 0% Z" scenario is narrower than that of the "50% Y - 50% Z" scenario, which is itself narrower than that of the "50% Y - 0% Z" and "50% Z - 0% Y" scenarios. Table 5 quantifies the differences in the sizes of the credibility intervals and present the reduction percentage in the credibility interval sizes of the "100% Y - 0% Z", "50% Y - 50% Z" and "50% Z - 0% Y" scenarios, compared to the "50% Y - 0% Z" scenario, for all the treatment efficacies at 6 days (third column) and at 12 days (fourth column) after pesticide spray. A positive (negative) value corresponds to an increase (decrease) of the credibility interval size compared to "50% Y - 0% Z". For example, the credibility interval size obtained for the efficacy of Movento at 6 days is 40.5% smaller with the "100% Y - 0% Z" scenario than with the "50% Y - 0% Z" scenario. Overall, the credibility interval sizes obtained with the "100% Y - 0% Z" scenario are 25 to 45% smaller than those obtained with the "50% Y - 0% Z" scenario. This indicates that the accuracy of the estimated treatment efficacies is improved by doubling the number of Y observations. Results show that the credibility intervals are often larger with "50% Z - 0% Y" than with "50% Y - 0% Z", suggesting that better esimtations are obtained using Y than Z when only a limited number of data is available, at least in this case study. Interestingly, the sizes of the credibility intervals are smaller by about 25% with "50% Y - 50% Z" compared to "50% Y - 0% Z", showing that the combination of Y and Z observations collected in different trials in beneficial and lead to a reduction of the uncertainty in the estimated treatment efficacy. The latter result illustrates how the estimation of treatment efficacy could be improved by combining trials with prevalence data and trials with intensity data.

@@

df_IC = rbind(df_resYhalf %>% mutate(Data = "50% Y - 0% Z"), 
              df_resYZ %>% mutate(Data = "50% Y - 50% Z"),
              df_resY %>% mutate(Data = "100% Y - 0% Z"),
              df_resZhalf %>% mutate(Data = "50% Z - 0% Y"))

@@

df_IC_6 = df_IC %>% mutate(Ordonnee = recode(Insecticide, !!!x)) %>%
                      mutate(Ordonnee = ifelse(Data == "100% Y - 0% Z", Ordonnee + y2, 
                                               ifelse(Data == "50% Y - 50% Z", Ordonnee + y1,
                                                      ifelse(Data == "50% Y - 0% Z", Ordonnee - y1, Ordonnee - y2))),
                             jour = "6")

# calcul du pourcentage de réduction de l'IC par rapport au scenario 50% Y - 0% Z
df_len = df_IC %>% select(Insecticide, name, value, Data) %>% pivot_wider(names_from =  c(name)) %>% as.data.frame %>%
    mutate(IC_length = b_sup - b_inf) %>% arrange(Insecticide)

df_remp = df_len %>% filter(Data == "50% Y - 0% Z") %>% select(Insecticide, IC_length)
x  = df_remp$IC_length;   names(x) = df_remp$Insecticide

@@

df_len6 = df_len %>% rename(Insecticide = Insecticide) %>% select(Insecticide, Data, `For efficacy at 6 days`) %>% 
    filter(Data != "50% Y - 0% Z") %>% arrange(Data) %>% mutate_if(is.numeric, round, digits = 1)
